### 2014-07-28 Studies/2012_Chicago/Report/Chicago_Fire.tex
Chicago: Minor edits throughout

\noindent Table~\ref{tab:fire_info} shows an overview of the timeline of events. A plan view of the structure and the location where the captain collapsed are shown in Fig.~(\ref{fig:simp_geom}).
\begin{table}
\centering
\captionof{table}{Abridged NOISH approximate fire event timeline}\label{tab:fire_info}
\begin{tabular}{cl}
\toprule[1.5pt]
Incident Time  &  Fire Behavior / Fireground Operation  \\
{[hh:mm:ss]}   &                                        \\
\midrule
\multirow{2}{*}{17:15:00} & \multirow{2}{*}{\parbox{10cm} {Reported working fire with heavy smoke coming from the front and rear of the structure's attic.}} \\ 
         & \\[.25cm] %blank rows exist for wrapping text. # of blanks is 1 less than number of rows
\multirow{2}{*}{17:16:00}    &  \multirow{2}{*}{\parbox{10cm} {Dispatch for a ``Still'' alarm - two engine companies, two truck companies, and a battalion chief.}} \\
         & \\[.25cm]
\multirow{2}{*}{17:20:00}    & \multirow{2}{*}{\parbox{10cm} {Attic windows break - heavy smoke out of front attic window and heavy smoke and fire out of rear.}} \\
         & \\[.25cm] 
17:21:00 & E123 captain and firefighter on 2nd floor with 1\sfrac{3}{4}-inch hoseline. \\[.25cm]
17:21:30 & 1st floor exterior porch door opened. \\[.25cm]
\multirow{2}{*}{17:23:00}    & \multirow{2}{*}{\parbox{10cm} {Heavy fire report in rear stairwell and enclosed porch. No response from E123 captain or firefighter.}} \\
         & \\[.25cm]
17:24:00    & Water applied to attic from the rear exterior position. \\[.25cm]
17:25:00    & SQ5 makes entry through front of structure. \\[.25cm]
17:27:00    & After hearing a Mayday call from SQ5, IC calls a Mayday. \\[.25cm]
17:29:00    & E123 captain carried out of structure. \\
\bottomrule[1.25pt]
\end{tabular}\par
\end{table}

@@

\begin{figure}[!ht]
\begin{tabular*}{\textwidth}{l@{\extracolsep{\fill}}r}
\includegraphics[width=\textwidth]{../Figures/1st_Floor_Metric} \\
\includegraphics[width=\textwidth]{../Figures/2nd_Floor_Metric}
\end{tabular*}
\caption[Plan view of 1st (top) and 2nd (bottom) floor of structure.]{Plan view of 1st (top) and 2nd (bottom) floor of structure. Stairs and key rooms are identified using measurements collected by NIST.}
\label{fig:simp_geom}
\end{figure}

@@

\begin{figure}[!ht]
\centering
\includegraphics[width=.60\textwidth]{../Figures/smv_exterior_mesh}
\caption[Front-right side of the structure within the computational domain (with multiple meshes).]{Front-right side of the structure within the computational domain (with multiple meshes). Each of the 16 meshes is 3~m (9.8~ft) by 10~m (33~ft) by 5~m (16.4~ft).}
\label{fig:mult_mesh}
\end{figure}

@@

\begin{table}
\centering
\captionof{table}{Fire incident and simulation event timeline}\label{tab:firesim}
\begin{tabular}{ccl}
\toprule[1.5pt]
Incident Time  &  Simulation Time  &  Fire Behavior / Fireground Operation  \\
{[hh:mm:ss]}   &  {[s]}            &                                        \\
\midrule
\multirow{2}{*}{17:15:00} &    & \multirow{2}{*}{\parbox{9cm} {Reported working fire with heavy smoke coming from the front and rear of the structure's attic.}} \\ 
         & & \\[.25cm] %blank rows exist for wrapping text. # of blanks is 1 less than number of rows
\multirow{2}{*}{17:16:00} & \multirow{2}{*}{}   &  \multirow{2}{*}{\parbox{9cm} {Dispatch for a ``Still'' alarm - two engine companies, two truck companies, and a battalion chief.}} \\
         & \\[.25cm]
         &  \multirow{2}{*} {0} &  \multirow{2}{*}{\parbox{9cm} {Simulation starts. Attic fire ignites. Initial ventilation through rear of structure. Front door open.}} \\
         & & \\[0.25cm]
\multirow{3}{*}{17:20:00} & \multirow{3}{*}{40}  & \multirow{3}{*}{\parbox{9cm} {Attic windows break - heavy smoke out of front attic window and heavy smoke and fire out of rear. Rear porch ignites in model.}} \\
         & & \\[.25cm]
         & & \\[.25cm]  
         & 45 & Ventilation through wall penetration on rear of structure. \\[.25cm] 
         & 55 & Ventilation through wall penetration on left side of structure. \\[.25cm] 
17:21:00 & 100 & E123 captain and firefighter on 2nd floor with 1\sfrac{3}{4}-inch hoseline. \\[.25cm]
\multirow{2}{*}{17:21:30} & \multirow{2}{*}{130} &  \multirow{2}{*}{\parbox{9cm}{1st floor exterior porch door opened. Porch fire ignites in FDS model.}} \\
        &  & \\[.25cm]
        & 135 & Additional ventilation through left side of structure. \\[.25cm] 
        & 160 & Top of interior door fails in FDS model. \\[.25cm] 
\multirow{2}{*}{17:23:00}    & \multirow{2}{*}{220} & \multirow{2}{*}{\parbox{9cm} {Heavy fire report in rear stairwell and enclosed porch. No response from E123 captain or firefighter.}} \\
         & & \\[.25cm]
17:24:00    &   & Water applied to attic from the rear exterior position. \\[.25cm]
17:25:00    &   & SQ5 makes entry through front of structure. \\[.25cm]
\multirow{2}{*}{17:27:00}    & \multirow{2}{*}{300}  & \multirow{2}{*}{\parbox{9cm} {After hearing a Mayday call from SQ5, IC calls a Mayday. Simulation ends.}}\\
         & &  \\[.25cm]
17:29:00    &   & E123 captain carried out of structure. \\
\bottomrule[1.25pt]
\end{tabular}\par
\end{table}

@@

\begin{figure}[!ht]
\centering
\includegraphics[width=.80\textwidth]{../Figures/Chicago_Fire_HRR}
\caption[HRR versus time from FDS simulations of the West 50th St. fire.]{HRR versus time from FDS simulations of the West 50th St. fire. The first vertical line indicates when the back door failed, and the second vertical line indicates when the interior hallway door failed.}
\label{fig:hrr}
\end{figure}

@@

Figure~(\ref{fig:temp_129s}) shows temperatures of at least 260~$^{\circ}$C (500~$^{\circ}$F) in the second floor rear porch and rear portion of the attic. High temperatures, between 80~$^{\circ}$C (185~$^{\circ}$F) and 200~$^{\circ}$C (390~$^{\circ}$F), also extend down to the rear entrance on the 1st floor. Before the interior door fails (Fig.~(\ref{fig:temp_159s})), the inflow of ambient air from the open rear door reduces the temperatures near the interior doorway, but at the same time the HRR increases because of the additional ventilation, as shown in Fig.~(\ref{fig:hrr}). As shown in Fig.~(\ref{fig:temp_161s}), after the second floor interior door fails, the high temperature gas that filled the rear porch flows in the direction of lower pressure (i.e., towards the front door). Note that the conditions in the attic do not appear to change.
\begin{figure}[!ht]
\includegraphics[width=.675\textwidth]{../Figures/west_50th_baseline_161}
\input{../Scripts/colorbar_std} 
\input{../Scripts/colorbar_metric}
\caption{FDS simulated temperature contours, 1~s after interior door failure.}
\label{fig:temp_161s}
\end{figure}

@@

\begin{figure}[!ht]
\centering
\includegraphics[width=.7\textwidth]{../Figures/ChicagoFlow}
\caption{Flow path on the second floor after the interior door failure.}
\label{fig:flowpath_1}
\end{figure}

@@

Prior to the interior door failure, the elevated temperatures on the second floor were limited to the rear porch, as shown in Fig.~(\ref{fig:temp_top_159}). The temperatures were greatest at the back-left corner where there is sufficient oxygen due to the section of porch wall that had burned away and allowed additional oxygen to mix with the fuel-rich fire gases. Five seconds after the interior door failed, hot gases traveled the length of the interior hallway and approached the doorway into the kitchen, which was approximately 4.5~m (14.8~ft) away from the source fire, as shown in Fig.~(\ref{fig:temp_top_165}). Within one minute of the interior door failure, the hallway temperatures were approximately at or above 260~$^{\circ}$C (500~$^{\circ}$F) throughout the hallway, as shown in Fig.~(\ref{fig:temp_top_220}). These results indicate that the high temperature gases seen 5~s after failure were not just a burst but rather the front of a sustained exhaust gas flow from the fire burning in the enclosed rear porch. The conditions in this hallway changed from tenable to high-hazard very rapidly following the door failure. Figure~(\ref{fig:temp_top_220}) also shows that the temperatures in the kitchen were at or below 90~$^{\circ}$C (194~$^{\circ}$F), and the temperatures at the top of the stairs were at or below 70~$^{\circ}$C (158~$^{\circ}$F). At these locations within the structure, the exhaust gases in the flow path were cooled and diluted by the inflowing ambient gas. These simulated temperatures are consistent with the post-incident conditions that were documented in the kitchen and second floor stairwell.
\begin{figure}[!ht]
\centering
\includegraphics[width=.68\textwidth]{../Figures/west_50th_baseline_top_159_6ft}
\input{../Scripts/colorbar_std} 
\input{../Scripts/colorbar_metric}
\caption{Temperature contours on the second floor, 1~s prior to door failure, 1.83~m (6~ft) above the floor.}
\label{fig:temp_top_159}
\end{figure}

\begin{figure}[!ht]
\centering
\includegraphics[width=.68\textwidth]{../Figures/west_50th_baseline_top_165_6ft}
\input{../Scripts/colorbar_std} 
\input{../Scripts/colorbar_metric}
\caption{Temperature contours on the second floor, 5~s after door failure, 1.83~m (6~ft) above the floor.}
\label{fig:temp_top_165}
\end{figure}

\begin{figure}[!ht]
\centering
\includegraphics[width=.68\textwidth]{../Figures/west_50th_baseline_top_220_6ft}
\input{../Scripts/colorbar_std} 
\input{../Scripts/colorbar_metric}
\caption{Temperature contours on the second floor, 60~s after door failure, 1.83~m (6~ft) above the floor.}
\label{fig:temp_top_220}
\end{figure}

@@

\begin{figure}[!ht]
\centering
\includegraphics[width=1\textwidth]{../Figures/50th_Place_1st_Floor}
\caption[Dimensioned drawing of 1st floor.]{Dimensioned drawing of 1st floor. The measurements are accurate to within 15~cm (6~in).}
\label{fig:first_floor}
\end{figure}

\begin{figure}[!ht]
\centering
\includegraphics[width=1\textwidth]{../Figures/50th_Place_2nd_Floor}
\caption[Dimensioned drawing of 2nd floor.]{Dimensioned drawing of 2nd floor. The measurements are accurate to within 15~cm (6~in).}
\label{fig:second_floor}
\end{figure}

>>>>

\noindent Table~\ref{tab:fire_info} shows an overview of the timeline of events. A plan view of the structure and the location where the captain collapsed are shown in Fig.~(\ref{fig:simp_geom}).
\begin{table}
\centering
\captionof{table}{Abridged NOISH approximate fire event timeline}\label{tab:fire_info}
\begin{tabular}{cl}
\toprule[1.5pt]
Incident Time  &  Fire Behavior / Fireground Operation  \\
{[hh:mm:ss]}   &                                        \\
\midrule
\multirow{2}{*}{17:15:00} & \multirow{2}{*}{\parbox{10cm} {Reported working fire with heavy smoke coming from the front and rear of the structure's attic.}} \\ 
         & \\[.25cm] %blank rows exist for wrapping text. # of blanks is 1 less than number of rows
\multirow{2}{*}{17:16:00}    &  \multirow{2}{*}{\parbox{10cm} {Dispatch for a ``Still'' alarm - two engine companies, two truck companies, and a battalion chief.}} \\
         & \\[.25cm]
\multirow{2}{*}{17:20:00}    & \multirow{2}{*}{\parbox{10cm} {Attic windows break - heavy smoke out of front attic window and heavy smoke and fire out of rear.}} \\
         & \\[.25cm] 
17:21:00 & E123 captain and firefighter on second floor with 1\sfrac{3}{4}-inch hoseline. \\[.25cm]
17:21:30 & First floor exterior porch door opened. \\[.25cm]
\multirow{2}{*}{17:23:00}    & \multirow{2}{*}{\parbox{10cm} {Heavy fire report in rear stairwell and enclosed porch. No response from E123 captain or firefighter.}} \\
         & \\[.25cm]
17:24:00    & Water applied to attic from the rear exterior position. \\[.25cm]
17:25:00    & SQ5 makes entry through front of structure. \\[.25cm]
17:27:00    & After hearing a Mayday call from SQ5, IC calls a Mayday. \\[.25cm]
17:29:00    & E123 captain carried out of structure. \\
\bottomrule[1.25pt]
\end{tabular}\par
\end{table}

@@

\begin{figure}[!ht]
\begin{tabular*}{\textwidth}{l@{\extracolsep{\fill}}r}
\includegraphics[width=\textwidth]{../Figures/1st_Floor_Metric} \\
\includegraphics[width=\textwidth]{../Figures/2nd_Floor_Metric}
\end{tabular*}
\caption[Plan view of first (top) and second (bottom) floor of structure.]{Plan view of first (top) and second (bottom) floor of structure. Stairs and key rooms are identified using measurements collected by NIST.}
\label{fig:simp_geom}
\end{figure}

@@

\begin{figure}[!ht]
\centering
\includegraphics[width=.60\textwidth]{../Figures/smv_exterior_mesh}
\caption[Front-right side of the structure within the computational domain.]{Front-right side of the structure within the computational domain (with multiple meshes). Each of the 16 meshes is 3~m (9.8~ft) by 10~m (33~ft) by 5~m (16.4~ft).}
\label{fig:mult_mesh}
\end{figure}

@@

\begin{table}
\centering
\captionof{table}{Fire incident and simulation event timeline}\label{tab:firesim}
\begin{tabular}{ccl}
\toprule[1.5pt]
Incident Time  &  Simulation Time  &  Fire Behavior / Fireground Operation  \\
{[hh:mm:ss]}   &  {[s]}            &                                        \\
\midrule
\multirow{2}{*}{17:15:00} &    & \multirow{2}{*}{\parbox{9cm} {Reported working fire with heavy smoke coming from the front and rear of the structure's attic.}} \\ 
         & & \\[.25cm] %blank rows exist for wrapping text. # of blanks is 1 less than number of rows
\multirow{2}{*}{17:16:00} & \multirow{2}{*}{}   &  \multirow{2}{*}{\parbox{9cm} {Dispatch for a ``Still'' alarm - two engine companies, two truck companies, and a battalion chief.}} \\
         & \\[.25cm]
         &  \multirow{2}{*} {0} &  \multirow{2}{*}{\parbox{9cm} {Simulation starts. Attic fire ignites. Initial ventilation through rear of structure. Front door open.}} \\
         & & \\[0.25cm]
\multirow{3}{*}{17:20:00} & \multirow{3}{*}{40}  & \multirow{3}{*}{\parbox{9cm} {Attic windows break - heavy smoke out of front attic window and heavy smoke and fire out of rear. Rear porch ignites in model.}} \\
         & & \\[.25cm]
         & & \\[.25cm]  
         & 45 & Ventilation through wall penetration on rear of structure. \\[.25cm] 
         & 55 & Ventilation through wall penetration on left side of structure. \\[.25cm] 
17:21:00 & 100 & E123 captain and firefighter on second floor with 1\sfrac{3}{4}-inch hoseline. \\[.25cm]
\multirow{2}{*}{17:21:30} & \multirow{2}{*}{130} &  \multirow{2}{*}{\parbox{9cm}{First floor exterior porch door opened. Porch fire ignites in FDS model.}} \\
        &  & \\[.25cm]
        & 135 & Additional ventilation through left side of structure. \\[.25cm] 
        & 160 & Top of interior door fails in FDS model. \\[.25cm] 
\multirow{2}{*}{17:23:00}    & \multirow{2}{*}{220} & \multirow{2}{*}{\parbox{9cm} {Heavy fire report in rear stairwell and enclosed porch. No response from E123 captain or firefighter.}} \\
         & & \\[.25cm]
17:24:00    &   & Water applied to attic from the rear exterior position. \\[.25cm]
17:25:00    &   & SQ5 makes entry through front of structure. \\[.25cm]
\multirow{2}{*}{17:27:00}    & \multirow{2}{*}{300}  & \multirow{2}{*}{\parbox{9cm} {After hearing a Mayday call from SQ5, IC calls a Mayday. Simulation ends.}}\\
         & &  \\[.25cm]
17:29:00    &   & E123 captain carried out of structure. \\
\bottomrule[1.25pt]
\end{tabular}\par
\end{table}

@@

\begin{figure}[!ht]
\centering
\includegraphics[width=.80\textwidth]{../Figures/Chicago_Fire_HRR}
\caption[HRR versus time from the FDS simulations.]{HRR versus time from the FDS simulations. The first vertical line indicates the time at which the back door failed, and the second vertical line indicates the time at which the interior hallway door failed.}
\label{fig:hrr}
\end{figure}

@@

Figure~(\ref{fig:temp_129s}) shows temperatures of at least 260~$^{\circ}$C (500~$^{\circ}$F) in the second floor rear porch and rear portion of the attic. High temperatures, between 80~$^{\circ}$C (185~$^{\circ}$F) and 200~$^{\circ}$C (390~$^{\circ}$F), also extend down to the rear entrance on the first floor. Before the interior door fails (Fig.~(\ref{fig:temp_159s})), the inflow of ambient air from the open rear door reduces the temperatures near the interior doorway, but at the same time the HRR increases because of the additional ventilation, as shown in Fig.~(\ref{fig:hrr}). As shown in Fig.~(\ref{fig:temp_161s}), after the second floor interior door fails, the high temperature gas that filled the rear porch flows in the direction of lower pressure (i.e., towards the front door). Note that the conditions in the attic do not appear to change.
\begin{figure}[!ht]
\includegraphics[width=.675\textwidth]{../Figures/west_50th_baseline_161}
\input{../Scripts/colorbar_std} 
\input{../Scripts/colorbar_metric}
\caption{FDS simulated temperature contours, 1~s after interior door failure.}
\label{fig:temp_161s}
\end{figure}

@@

\begin{figure}[!ht]
\centering
\includegraphics[width=.7\textwidth]{../Figures/ChicagoFlow}
\caption{Flow path on the second floor after interior door failure.}
\label{fig:flowpath_1}
\end{figure}

@@

Prior to the interior door failure, the elevated temperatures on the second floor were limited to the rear porch, as shown in Fig.~(\ref{fig:temp_top_159}). The temperatures were greatest at the back-left corner where there is sufficient oxygen due to the section of porch wall that had burned away and allowed additional oxygen to mix with the fuel-rich fire gases. Five seconds after the interior door failed, hot gases traveled the length of the interior hallway and approached the doorway into the kitchen, which was approximately 4.5~m (14.8~ft) away from the source fire, as shown in Fig.~(\ref{fig:temp_top_165}). Within one minute of the interior door failure, the hallway temperatures were approximately at or above 260~$^{\circ}$C (500~$^{\circ}$F) throughout the hallway, as shown in Fig.~(\ref{fig:temp_top_220}). These results indicate that the high temperature gases seen 5~s after failure were not just a burst but rather the front of a sustained exhaust gas flow from the fire burning in the enclosed rear porch. The conditions in this hallway changed from tenable to high-hazard very rapidly following the door failure. Figure~(\ref{fig:temp_top_220}) also shows that the temperatures in the kitchen were at or below 90~$^{\circ}$C (194~$^{\circ}$F), and the temperatures at the top of the stairs were at or below 70~$^{\circ}$C (158~$^{\circ}$F). At these locations within the structure, the exhaust gases in the flow path were cooled and diluted by the inflowing ambient gas. These simulated temperatures are consistent with the post-incident conditions that were documented in the kitchen and second floor stairwell.
\begin{figure}[!ht]
\centering
\includegraphics[width=.68\textwidth]{../Figures/west_50th_baseline_top_159_6ft}
\input{../Scripts/colorbar_std} 
\input{../Scripts/colorbar_metric}
\caption[Temperature contours on the second floor, 1~s prior to interior door failure.]
{Temperature contours on the second floor, 1~s prior to interior door failure, 1.83~m (6~ft) above the floor.}
\label{fig:temp_top_159}
\end{figure}

\begin{figure}[!ht]
\centering
\includegraphics[width=.68\textwidth]{../Figures/west_50th_baseline_top_165_6ft}
\input{../Scripts/colorbar_std} 
\input{../Scripts/colorbar_metric}
\caption[Temperature contours on the second floor, 5~s after interior door failure.]
{Temperature contours on the second floor, 5~s after interior door failure, 1.83~m (6~ft) above the floor.}
\label{fig:temp_top_165}
\end{figure}

\begin{figure}[!ht]
\centering
\includegraphics[width=.68\textwidth]{../Figures/west_50th_baseline_top_220_6ft}
\input{../Scripts/colorbar_std} 
\input{../Scripts/colorbar_metric}
\caption[Temperature contours on the second floor, 60~s after interior door failure.]
{Temperature contours on the second floor, 60~s after interior door failure, 1.83~m (6~ft) above the floor.}
\label{fig:temp_top_220}
\end{figure}

@@

\begin{figure}[!ht]
\centering
\includegraphics[width=1\textwidth]{../Figures/50th_Place_1st_Floor}
\caption[Dimensioned drawing of first floor.]{Dimensioned drawing of first floor. The measurements are accurate to within 15~cm (6~in).}
\label{fig:first_floor}
\end{figure}

\begin{figure}[!ht]
\centering
\includegraphics[width=1\textwidth]{../Figures/50th_Place_2nd_Floor}
\caption[Dimensioned drawing of second floor.]{Dimensioned drawing of second floor. The measurements are accurate to within 15~cm (6~in).}
\label{fig:second_floor}
\end{figure}
